On the probability of finding a nonphysical solution through shadowing\tnoteref{1}

### 2020-12-08 main_els.tex
Add new elsarticle class.

%%%%%%%%%%%%%%%%%%%%%%% file template.tex %%%%%%%%%%%%%%%%%%%%%%%%%
%
% This is a general template file for the LaTeX package SVJour3
% for Springer journals.          Springer Heidelberg 2010/09/16
%
% Copy it to a new file with a new name and use it as the basis
% for your article. Delete % signs as needed.
%
% This template includes a few options for different layouts and
% content for various journals. Please consult a previous issue of
% your journal as needed.
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%
% First comes an example EPS file -- just ignore it and
% proceed on the \documentclass line
% your LaTeX will extract the file if required
%\begin{filecontents*}{example.eps}
%!PS-Adobe-3.0 EPSF-3.0
%%BoundingBox: 19 19 221 221
%%CreationDate: Mon Sep 29 1997
%%Creator: programmed by hand (JK)
%%EndComments
%gsave
%newpath
%  20 20 moveto
%  20 220 lineto
%  220 220 lineto
%  220 20 lineto
%closepath
%2 setlinewidth
%gsave
%  .4 setgray fill
%grestore
%stroke
%grestore
%\end{filecontents*}
%
\RequirePackage{fix-cm}
%
%\documentclass{svjour3}                     % onecolumn (standard format)
%\documentclass[smallcondensed]{svjour3}     % onecolumn (ditto)
\documentclass[smallextended]{svjour3}       % onecolumn (second format)
%\documentclass[twocolumn]{svjour3}          % twocolumn
%
\smartqed  % flush right qed marks, e.g. at end of proof
%
% \usepackage{mathptmx}      % use Times fonts if available on your TeX system
%
% insert here the call for the packages your document requires
%\usepackage{latexsym}
% etc.
%
% please place your own definitions here and don't use \def but
% \newcommand{}{}
%
% Insert the name of "your journal" with
% \journalname{myjournal}
\usepackage[utf8]{inputenc}
\usepackage{graphicx, float}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{mathtools,hyperref}
\usepackage{xcolor,verbatim}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\newcommand{\nisha}[1]{{\color{darkgreen} #1}}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}
\begin{document}

\title{On the probability of finding a nonphysical solution through shadowing\footnote{Preprint available at \cite{preprint}}}
\date{June 2020}

%\titlerunning{Short form of title}        % if too long for running head

\author{Nisha Chandramoorthy \and Qiqi Wang}

%\authorrunning{Short form of author list} % if too long for running head
\institute{N. Chandramoorthy, Corresponding Author \at
              Mechanical Engineering and Center for Computational Science and Engineering \\
              77 Massachusetts Avenue, Cambridge MA
               \email{nishac@mit.edu}           %  \\
%             \emph{Present address:} of F. Author  %  if needed
           \and
           Q. Wang \at
              Department of Aeronautics and Astronautics and Center for Computational Science and Engineering \\
              77 Massachusetts Avenue, Cambridge MA \\
              \email{qiqi@mit.edu}
}
%\date{Received: date / Accepted: date}
% The correct dates will be entered by the editor

\maketitle

\begin{abstract}
This paper proves that shadowing solutions can be almost surely nonphysical.
This finding invalidates the argument that small perturbations in a chaotic system can only have a small impact on its statistical behavior.
This theoretical finding has implications for many applications
in which chaotic mechanics plays an important role.
It suggests, for example, that we can control the climate through subtle perturbations. It also suggests that numerical simulations of chaotic dynamics, such as turbulent flows and global atmosphere and ocean circulation, may fail to predict the true long-term or statistical behavior.
\keywords{Climate \and Chaotic systems \and Numerical simulations \and Turbulence \and Shadowing sensitivity analysis}
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

\maketitle
\section{Introduction}
In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models. We construct numerical and analytical examples that illustrate this feature of chaotic dynamics. The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere. This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect. Thus, the main conjecture of this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations. These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions. In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
\begin{itemize}
    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we eliminate model uncertainties stemming from imperfect modeling of the underlying physics, and statistical noise due to finite-time averaging;
    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods for sensitivity computation;
    \item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior. This means that the Lorenzian butterfly can significantly alter the climate (statistical/long-term average) of Texas, and not just its weather (short-term events).
\end{itemize}
The paper is organized as follows. In section \ref{sec:shadowingIntro}, we introduce the dynamical systems concept of shadowing and its applications in the context of chaotic numerical simulations. In section \ref{sec:nonphysical}, we define the notion of physicality and give examples to illustrate the different ways in which to obtain nonphysical solutions. Section \ref{sec:shadowingnonphysical} is dedicated to several examples of perturbed tent maps that have nonphysical shadowing solutions. In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change the statistics significantly. The supplementary material, section \ref{sec:supp}, contains the numerical procedure to generate the probability distributions of the shadowing solutions.

@@

\bibliographystyle{amm}

@@

But, as mentioned in the main text, even a small perturbation -- very small $s$ -- is sufficient to trigger a remarkable variation in the statistics. This is because this particular construction transfers 
the nonuniformities in the probability distribution at larger scales to smaller scales, with increasing $n.$ The nonuniformity at the largest 
scale of half intervals is retained at higher values of $n.$ That is,
the probability of visiting the interval $[0,1]$ is the same: $(1-s)$, at all $n$. However, with increasing $n,$ the nonuniformity emerges within smaller subintervals, due to the construction of the map that relies on repeating the behavior at larger scales at smaller scales (Eq. \ref{eqn:frequency}). This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMap}, bottom-right). The scripts that generate the plucked tent map and its 
physical probability distributions can be found under \verb+tent_sens_stat+. 
\end{document}

>>>>

%% 
%% Copyright 2007-2020 Elsevier Ltd
%% 
%% This file is part of the 'Elsarticle Bundle'.
%% ---------------------------------------------
%% 
%% It may be distributed under the conditions of the LaTeX Project Public
%% License, either version 1.2 of this license or (at your option) any
%% later version.  The latest version of this license is in
%%    http://www.latex-project.org/lppl.txt
%% and version 1.2 or later is part of all distributions of LaTeX
%% version 1999/12/01 or later.
%% 
%% The list of all files belonging to the 'Elsarticle Bundle' is
%% given in the file `manifest.txt'.
%%

%% Template article for Elsevier's document class `elsarticle'
%% with numbered style bibliographic references
%% SP 2008/03/01
%%
%% 
%%
%% $Id: elsarticle-template-num.tex 190 2020-11-23 11:12:32Z rishi $
%%
%%
\documentclass[preprint,12pt]{elsarticle}

%% Use the option review to obtain double line spacing
%% \documentclass[authoryear,preprint,review,12pt]{elsarticle}

%% Use the options 1p,twocolumn; 3p; 3p,twocolumn; 5p; or 5p,twocolumn
%% for a journal layout:
%% \documentclass[final,1p,times]{elsarticle}
%% \documentclass[final,1p,times,twocolumn]{elsarticle}
%% \documentclass[final,3p,times]{elsarticle}
%% \documentclass[final,3p,times,twocolumn]{elsarticle}
%% \documentclass[final,5p,times]{elsarticle}
%% \documentclass[final,5p,times,twocolumn]{elsarticle}

%% For including figures, graphicx.sty has been loaded in
%% elsarticle.cls. If you prefer to use the old commands
%% please give \usepackage{epsfig}

%% The amssymb package provides various useful mathematical symbols
\usepackage[utf8]{inputenc}
\usepackage{graphicx, float}
\usepackage{amsmath, amssymb, amsfonts}
\usepackage{mathtools}
\usepackage{xcolor,verbatim}
\definecolor{darkgreen}{rgb}{0.0, 0.5, 0.0}
\newcommand{\nisha}[1]{{\color{darkgreen} #1}}
\DeclarePairedDelimiter{\floor}{\lfloor}{\rfloor}

%% The amsthm package provides extended theorem environments
%% \usepackage{amsthm}

%% The lineno packages adds line numbers. Start line numbering with
%% \begin{linenumbers}, end it with \end{linenumbers}. Or switch it on
%% for the whole article with \linenumbers.
%% \usepackage{lineno}

\journal{Journal of Computational Physics}

\begin{document}

\begin{frontmatter}

%% Title, authors and addresses

%% use the tnoteref command within \title for footnotes;
%% use the tnotetext command for theassociated footnote;
%% use the fnref command within \author or \address for footnotes;
%% use the fntext command for theassociated footnote;
%% use the corref command within \author for corresponding author footnotes;
%% use the cortext command for theassociated footnote;
%% use the ead command for the email address,
%% and the form \ead[url] for the home page:
%% \title{Title\tnoteref{label1}}
%% \tnotetext[label1]{}
%% \author{Name\corref{cor1}\fnref{label2}}
%% \ead{email address}
%% \ead[url]{home page}
%% \fntext[label2]{}
%% \cortext[cor1]{}
%% \affiliation{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%% \fntext[label3]{}

		\title{On the probability of finding a nonphysical solution through shadowing\tnoteref{1}}
		\tnotetext[1]{Preprint available at \cite{preprint}}

%% use optional labels to link authors explicitly to addresses:
%% \author[label1,label2]{}
%% \affiliation[label1]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}
%%
%% \affiliation[label2]{organization={},
%%             addressline={},
%%             city={},
%%             postcode={},
%%             state={},
%%             country={}}

\author[N,ccse]{Nisha Chandramoorthy\corref{cor}}
\cortext[cor]{Corresponding author}
\ead{nishac@mit.edu}
\affiliation[ccse]{organization={Center for Computational Science and Engineering, Massachusetts Institute of Technology},%Department and Organization
            addressline={77 Massachusetts Avenue}, 
            city={Cambridge},
            postcode={02139}, 
            state={MA},
            country={USA}}
\author[Q,ccse]{Qiqi Wang}
\ead{qiqi@mit.edu}
\affiliation[N]{Department of Mechanical Engineering}
\affiliation[Q]{Department of Aeronautics and Astronautics}
\begin{abstract}
This paper proves that shadowing solutions can be almost surely nonphysical.
This finding invalidates the argument that small perturbations in a chaotic system can only have a small impact on its statistical behavior.
This theoretical finding has implications for many applications
in which chaotic mechanics plays an important role.
It suggests, for example, that we can control the climate through subtle perturbations. It also suggests that numerical simulations of chaotic dynamics, such as turbulent flows and global atmosphere and ocean circulation, may fail to predict the true long-term or statistical behavior.
% \PACS{PACS code1 \and PACS code2 \and more}
% \subclass{MSC code1 \and MSC code2 \and more}
\end{abstract}

%%Research highlights
\begin{highlights}
\item Research highlight 1
\item Research highlight 2
\end{highlights}

\begin{keyword}
Chaotic dynamics \sep numerical simulations \sep scientific computing 
\sep shadowing sensitivity analysis %% keywords here, in the form: keyword \sep keyword

%% PACS codes here, in the form: \PACS code \sep code

%% MSC codes here, in the form: \MSC code \sep code
%% or \MSC[2008] code \sep code (2000 is the default)

\end{keyword}

\end{frontmatter}

%% \linenumbers

\section{Introduction}
In this paper, we show that a large mismatch is possible in the statistical or long-time averages measured in a pair of chaotic solutions to slightly different models. We construct numerical and analytical examples that illustrate this feature of chaotic dynamics. The analysis of the long-term behavior of {\em shadowing} solutions in the majority of the chaotic models constructed in this paper reveals that shadowing solutions are almost surely {\em nonphysical}: they do not represent the behavior of a physically observed solution starting almost everywhere. This also leads us to the conclusion that shadowing-based sensitivities \cite{qiqi-lss} of long-time averages to parameter changes can be incorrect. Thus, the main conjecture of this paper urges us to reconsider the use of chaotic numerical simulations, and statistical sensitivity analysis on these simulations. These implications are particularly important for climate studies, where chaotic models are widely used to make long-term predictions. In particular, three implications of the evidence in this paper may be stated, the third of which harnesses the statistical inconsistency between slightly different models:
\begin{itemize}
    \item we cannot trust the long-time averages generated by numerical simulations of chaotic processes, even when we eliminate model uncertainties stemming from imperfect modeling of the underlying physics, and statistical noise due to finite-time averaging;
    \item we cannot trust the sensitivities of long-time averages or statistical averages to small parameter changes computed by using shadowing-based methods for sensitivity computation;
	\item it is possible for two models separated by tiny changes in parameter values to have drastically different statistical behavior. This means that the Lorenzian butterfly (a small perturbation) can significantly alter the climate of Texas (statistical/long-term average), and not just its weather (short-term events).
\end{itemize}
The paper is organized as follows. In section \ref{sec:shadowingIntro}, we introduce the dynamical systems concept of shadowing and its applications in the context of chaotic numerical simulations. In section \ref{sec:nonphysical}, we define the notion of physicality and give examples to illustrate the different ways in which to obtain nonphysical solutions. Section \ref{sec:shadowingnonphysical} is dedicated to several examples of perturbed tent maps that have nonphysical shadowing solutions. In the discussion in section \ref{sec:discussion}, we illustrate with an example that a small parameter perturbation can change the statistics significantly. The supplementary material, section \ref{sec:supp}, contains the numerical procedure to generate the probability distributions of the shadowing solutions.

@@

\bibliographystyle{elsarticle-num} 

@@

But, as mentioned in the main text, even a small perturbation -- very small $s$ -- is sufficient to trigger a remarkable variation in the statistics. This is because this particular construction transfers 
the nonuniformities in the probability distribution at larger scales to smaller scales, with increasing $n.$ The nonuniformity at the largest 
scale of half intervals is retained at higher values of $n.$ That is,
the probability of visiting the interval $[0,1]$ is the same: $(1-s)$, at all $n$. However, with increasing $n,$ the nonuniformity emerges within smaller subintervals, due to the construction of the map that relies on repeating the behavior at larger scales at smaller scales (Eq. \ref{eqn:frequency}). This repetitive construction is responsible for the apparent fractal structure of the probability distribution (Figure \ref{fig:pluckedTentMap}, bottom-right). The scripts that generate the plucked tent map and its 
physical probability distributions can be found under \verb+tent_sens_stat+.

\end{document}
\endinput
%%
